Unknown

### 2024-04-04 ms/ms.qmd
drafted light response cleaning methods and table

### Cleaning RH curves
n_rh_curve_raw = 2400 # export from adaptive-amphistomy/r/07_process-licor.R
n_points_per_rh_curves_raw = 100 # export from adaptive-amphistomy/r/07_process-licor.R
ca_min = aa_args$ca_min
ca_max = aa_args$ca_max
n_min_points_per_rh_curve = aa_args$n_min_points_per_rh_curve
point_outlier_threshold = aa_args$point_outlier_threshold
aa_outlier_threshold = aa_args$aa_outlier_threshold
thinning_interval = aa_args$thinning_interval

loo_version = "2.7.0"
```

@@

The raw data set consisted of `r prettyNum(aa_stats$n_rh_curve0)` \agcurve{} curves with an average of `r formatC(aa_stats$n_point_per_rh_curve0, digits = 3)` points per curve. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps \autoref{tbl-cleaning1}.

@@

n_curve_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_rh_curve[1-9]$")])

@@

*Procedure*: We fit provisional quadratic regressions for each curve using ordinary least squares with the `lm()` function in *R*. We sequentially removed data points with an absolute external studentized residual greater than `r point_outlier_threshold` until none remained.

@@

WORKING HERE!

The raw data set consisted of `r prettyNum(aa_stats$n_aq_curve0)` \aqcurve{} curves with an average of `r formatC(aa_stats$n_point_per_aq_curve0, digits = 3)`. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps \autoref{tbl-cleaning2}.

```{r, cleaning2, eval=FALSE,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-cleaning2
#| tbl-cap: "Six sequential steps for cleaning \\agcurve{} curves. The rationale and procedure for each step are described in the text. The rightmost columns summarize the number of curves and mean number of points per curve remaining after each step. For reference, there are four possible \\agcurve{} curves per replicate: all combinations of leaf type (amphi or pseudohypo) and light intensity (high or low)."

n_curve_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_rh_curve[0-9]$")])

n_point_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_point_per_rh_curve[0-9]$")])

@@

@@

- models fit in Stan/brms
- we chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was greater than $10^3$.

@@

We tested competing predictions through a combination of model selection and parameter estimation from the posterior of selected models. We started with a base model that includes fixed effects of light intensity and light treatment on \aax, a random effect of accession on \aax, and a linear regression of PAR on accession-level \aax. We next compared the fit of the base model to a more complex with an interaction between light intensity and light treatment using LOOIC with the \textit{R} package **loo** version `r loo_version` [@vehtari_practical_2017]. These comparisons allowed us to test the main predictions of each hypothesis and for a reinforcing interactions between assimilatory and plastic hypotheses (tab:predictions). However, these models cannot test predictions when there is an interaction with the constitutive hypothesis, because this hypothesis posits accession-levels variation in responses to light intensity and/or light treatment. Therefore, we tested for significant variation among accessions in light intensity and light treatment by comparing LOOIC values of models with and without random interactions between these factors and accession.

>>>>

### Cleaning RH curves
n_rh_curve_raw = 2400 # export from adaptive-amphistomy/r/07_process-licor.R
n_points_per_rh_curves_raw = 100 # export from adaptive-amphistomy/r/07_process-licor.R
ca_min = aa_args$ca_min
ca_max = aa_args$ca_max
n_min_points_per_rh_curve = aa_args$n_min_points_per_rh_curve
point_outlier_threshold = aa_args$point_outlier_threshold
aa_outlier_threshold = aa_args$aa_outlier_threshold
thinning_interval = aa_args$thinning_interval
aq_r2_threshold = aa_args$aq_r2_threshold

@@

The raw data set consisted of `r prettyNum(aa_stats$n_rh_curve0, big.mark = ",")` \agcurve{} curves with an average of `r formatC(aa_stats$n_point_per_rh_curve0, digits = 3)` points per curve. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into six sequential steps \autoref{tbl-cleaning1}.

@@

n_curve_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_rh_curve[1-9]$")]) |>
  prettyNum(big.mark = ",")

@@

*Procedure*: We fit provisional quadratic regressions for each curve using ordinary least squares with the `lm()` function in *R*. We sequentially removed data points with an absolute external studentized residual $> `r point_outlier_threshold`$ until none remained.

@@

The raw data set consisted of `r prettyNum(aa_stats$n_aq_curve0, big.mark = ",")` \aqcurve{} curves with an average of `r formatC(aa_stats$n_point_per_aq_curve0, digits = 3)`. Manual curation of a data set this size in a principled, consistent manner is not feasible. Therefore, we automated data cleaning using custom *R* scripts. Cleaning is divided into two sequential steps \autoref{tbl-cleaning2}.

```{r, cleaning2,echo=FALSE,message=FALSE,warning=FALSE}
#| label: tbl-cleaning2
#| tbl-cap: "Two sequential steps for cleaning \\agcurve{} curves. The rationale and procedure for each step are described in the text. The rightmost columns summarize the number of curves and mean number of points per curve remaining after each step."

step_description_col = c(
  "1. remove outliers within each \\aqcurve{} curve",
  "2. remove \\aqcurve{} curves with poor fit"
)

n_curve_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_aq_curve[1-9]$")]) |>
  prettyNum(big.mark = ",")

n_point_col = unlist(aa_stats[str_detect(names(aa_stats), "^n_point_per_aq_curve[1-9]$")])

@@

#### Remove outliers within each \aqcurve{} curve

*Procedure*: We fit provisional nonrectangular hyperbola [@marshall_model_1980] to each \aqcurve{} curve using nonlinear regression with the `nlsLM()` function from the *R* package **minpack.lm** version `r packageVersion("minpack.lm")` [@Elzhov_minpack.lm_2023]. We sequentially removed data points with an absolute external studentized residual $> `r point_outlier_threshold`$ until none remained.

#### Remove \aqcurve{} curves with poor fit

*Rationale*: \aqcurve{} curves with a poor fit to the nonrectangular hyperbola most likely indicate systematic measurement error and/or the leaf was not fully acclimated to the chamber environment.

*Procedure*: As described above, we fit provisional nonrectangular hyperbola to each \aqcurve{} curve and calculated the model $r^2$. There was a clear break between typical curves and poorly fitting curves where $r^2 < `r aq_r2_threshold`$. We therefore removed \aqcurve{} curves with $r^2 < `r aq_r2_threshold`$.

@@

- models fit in Stan/brms
- we chose the number of chains, warmup and sampling iterations, and maximum treedepth so that parameter estimates converged ($\hat{R} < 1.01$ [@gelman_inference_1992]) and the effective sample size (ESS) for each parameter was $> 10^3$.

@@

We tested competing predictions through a combination of model selection and parameter estimation from the posterior of selected models. We started with a base model that includes fixed effects of light intensity and light treatment on \aax, a random effect of accession on \aax, and a linear regression of PAR on accession-level \aax. We next compared the fit of the base model to a more complex with an interaction between light intensity and light treatment using LOOIC with the \textit{R} package **loo** version `r packageVersion("loo")` [@vehtari_practical_2017]. These comparisons allowed us to test the main predictions of each hypothesis and for a reinforcing interactions between assimilatory and plastic hypotheses (tab:predictions). However, these models cannot test predictions when there is an interaction with the constitutive hypothesis, because this hypothesis posits accession-levels variation in responses to light intensity and/or light treatment. Therefore, we tested for significant variation among accessions in light intensity and light treatment by comparing LOOIC values of models with and without random interactions between these factors and accession.

### 2024-04-04 ms/ms.tex
drafted light response cleaning methods and table

The raw data set consisted of 2370 \agcurve{} curves with an average of
63.2. Manual curation of a data set this size in a principled,
consistent manner is not feasible. Therefore, we automated data cleaning
using custom \emph{R} scripts. Cleaning is divided into six sequential
steps \autoref{tbl-cleaning1}.

@@

\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{9cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
\toprule
Step: description & Number of curves & Number of points per curve\\
\midrule
\cellcolor{gray!10}{1. remove unreliable and unusable data points} & \cellcolor{gray!10}{2361} & \cellcolor{gray!10}{63.0}\\
2. remove hysteretic portion of \agcurve{} curves at low \gsw{} & 2360 & 59.2\\
\cellcolor{gray!10}{3. remove outliers within each \agcurve{} curve} & \cellcolor{gray!10}{2360} & \cellcolor{gray!10}{58.7}\\
4. remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves & 2268 & 58.5\\
\cellcolor{gray!10}{5. thin redundant data points within each \agcurve{} curve} & \cellcolor{gray!10}{2268} & \cellcolor{gray!10}{28.1}\\
\addlinespace
6. Trim extreme \aax{} values & 2210 & 28.0\\
\bottomrule
\end{tabular}

@@

\emph{Procedure}: We fit provisional quadratic regressions for each
curve using ordinary least squares with the \texttt{lm()} function in
\emph{R}. We sequentially removed data points with an absolute external
studentized residual greater than 3 until none remained.

@@

WORKING HERE!

The raw data set consisted of \aqcurve{} curves with an average of .
Manual curation of a data set this size in a principled, consistent
manner is not feasible. Therefore, we automated data cleaning using
custom \emph{R} scripts. Cleaning is divided into six sequential steps
\autoref{tbl-cleaning2}.

@@

\caption{\label{tbl-cleaning2}Six sequential steps for cleaning
\agcurve{} curves. The rationale and procedure for each step are
described in the text. The rightmost columns summarize the number of
curves and mean number of points per curve remaining after each step.
For reference, there are four possible \agcurve{} curves per replicate:
all combinations of leaf type (amphi or pseudohypo) and light intensity
(high or low).}

@@

@@

@@

\begin{itemize}
\tightlist
\item
  models fit in Stan/brms
\item
  we chose the number of chains, warmup and sampling iterations, and
  maximum treedepth so that parameter estimates converged
  (\(\hat{R} < 1.01\) (\emph{8})) and the effective sample size (ESS)
  for each parameter was greater than \(10^3\).
\end{itemize}

>>>>

The raw data set consisted of 2,370 \agcurve{} curves with an average of
63.2 points per curve. Manual curation of a data set this size in a
principled, consistent manner is not feasible. Therefore, we automated
data cleaning using custom \emph{R} scripts. Cleaning is divided into
six sequential steps \autoref{tbl-cleaning1}.

@@

\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{9cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
\toprule
Step: description & Number of curves & Number of points per curve\\
\midrule
\cellcolor{gray!10}{1. remove unreliable and unusable data points} & \cellcolor{gray!10}{2,361} & \cellcolor{gray!10}{63.0}\\
2. remove hysteretic portion of \agcurve{} curves at low \gsw{} & 2,360 & 59.2\\
\cellcolor{gray!10}{3. remove outliers within each \agcurve{} curve} & \cellcolor{gray!10}{2,360} & \cellcolor{gray!10}{58.7}\\
4. remove replicates with no overlap between amphi and pseudohypo \agcurve{} curves & 2,268 & 58.5\\
\cellcolor{gray!10}{5. thin redundant data points within each \agcurve{} curve} & \cellcolor{gray!10}{2,268} & \cellcolor{gray!10}{28.1}\\
\addlinespace
6. Trim extreme \aax{} values & 2,210 & 28.0\\
\bottomrule
\end{tabular}

@@

\emph{Procedure}: We fit provisional quadratic regressions for each
curve using ordinary least squares with the \texttt{lm()} function in
\emph{R}. We sequentially removed data points with an absolute external
studentized residual \(> 3\) until none remained.

@@

The raw data set consisted of 658 \aqcurve{} curves with an average of
19.4. Manual curation of a data set this size in a principled,
consistent manner is not feasible. Therefore, we automated data cleaning
using custom \emph{R} scripts. Cleaning is divided into two sequential
steps \autoref{tbl-cleaning2}.

@@

\caption{\label{tbl-cleaning2}Two sequential steps for cleaning
\agcurve{} curves. The rationale and procedure for each step are
described in the text. The rightmost columns summarize the number of
curves and mean number of points per curve remaining after each step.}

@@

\centering
\begin{tabular}{>{\raggedright\arraybackslash}p{9cm}>{\centering\arraybackslash}p{2.5cm}>{\centering\arraybackslash}p{2.5cm}}
\toprule
Step: description & Number of curves & Number of points per curve\\
\midrule
\cellcolor{gray!10}{1. remove outliers within each \aqcurve{} curve} & \cellcolor{gray!10}{658} & \cellcolor{gray!10}{19.1}\\
2. remove \aqcurve{} curves with poor fit & 652 & 19.1\\
\bottomrule
\end{tabular}

@@

\paragraph{\texorpdfstring{Remove outliers within each \aqcurve{}
curve}{Remove outliers within each  curve}}\label{remove-outliers-within-each-curve-1}

\emph{Procedure}: We fit provisional nonrectangular hyperbola
(\textbf{marshall\_model\_1980?}) to each \aqcurve{} curve using
nonlinear regression with the \texttt{nlsLM()} function from the
\emph{R} package \textbf{minpack.lm} version 1.2.4
(\textbf{Elzhov\_minpack.lm\_2023?}). We sequentially removed data
points with an absolute external studentized residual \(> 3\) until none
remained.

\paragraph{\texorpdfstring{Remove \aqcurve{} curves with poor
fit}{Remove  curves with poor fit}}\label{remove-curves-with-poor-fit}

\emph{Rationale}: \aqcurve{} curves with a poor fit to the
nonrectangular hyperbola most likely indicate systematic measurement
error and/or the leaf was not fully acclimated to the chamber
environment.

\emph{Procedure}: As described above, we fit provisional nonrectangular
hyperbola to each \aqcurve{} curve and calculated the model \(r^2\).
There was a clear break between typical curves and poorly fitting curves
where \(r^2 < 0.99\). We therefore removed \aqcurve{} curves with
\(r^2 < 0.99\).

@@

\begin{itemize}
\tightlist
\item
  models fit in Stan/brms
\item
  we chose the number of chains, warmup and sampling iterations, and
  maximum treedepth so that parameter estimates converged
  (\(\hat{R} < 1.01\) (\emph{8})) and the effective sample size (ESS)
  for each parameter was \(> 10^3\).
\end{itemize}
